Follow-up Samples

### 2017-01-20 submission/manuscript_outline_20161024.Rmd
abstract improvements

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between fecal bacterial microbiome and CRC.   The overall objective of our study was to investigate the changes in the bacterial microbiome after surgery in patients with lesion (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion.  We also wanted to investigate the use of the bacterial microbiome and Fecal Immunoglobulin Test (FIT) to build models which could classify individuals as having a lesion or as before surgery.     
**Results:** Adenoma individual's bacterial microbiome were more similar to their pre-surgery sample then those with carcinoma (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`).  Their change in FIT was also significantly different then those with carcinoma (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).  There was no significant difference in any indivdiual OTU between samples before and after surgery (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while the model to classify samples as before surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  The after surgery samples had a significant decrease in the model positive probability of a lesion or before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria (`r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`).   
**Conclusions:** Our data suggests that treatment not only significantly reduces the probability of having a colonic lesion within our models but also causes detectable changes in the bacterial microbiome.  Further surveillance of these individuals will enable us to determine whether models such as the ones we present could be used to predict recurrence of colorectal cancer.

>>>>

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between fecal bacterial microbiome and CRC.   The overall objective of our study was to investigate the changes in the bacterial microbiome after surgery in patients with lesion (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion.  We also wanted to investigate the use of the bacterial microbiome and Fecal Immunoglobulin Test (FIT) to build models which could classify individuals as having a lesion or as before surgery.     
**Results:** Adenoma individual's bacterial microbiome were more similar to their pre-surgery sample then those with carcinoma (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`).  Their change in FIT was also significantly different then those with carcinoma (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).  There was no significant difference in any indivdiual OTU between samples before and after surgery (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while the model to classify samples as before surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  There was a significant decrease in the model positive probability for the after surgery sample being either a lesion or before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria (`r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`).   
**Conclusions:** Our data suggests that treatment not only significantly reduces the probability of having a colonic lesion within our models but also causes detectable changes in the bacterial microbiome.  Further surveillance of these individuals will enable us to determine whether models such as the ones we present could be used to predict recurrence of colorectal cancer.

### 2017-01-20 submission/manuscript_outline_20161024.tex
add second SRA accession
abstract improvements

\textbf{Background:} Colorectal cancer (CRC) continues to be a worldwide
health problem with early detection being used as a key component in
mitigating deaths due to the disease. Previous research suggests a link
between fecal bacterial microbiome and CRC. The overall objective of our
study was to investigate the changes in the bacterial microbiome after
surgery in patients with lesion (i.e.~adenoma or carcinoma).
Specifically, we wanted to identify what within the community was
different within those undergoing surgical removal of lesion. We also
wanted to investigate the use of the bacterial microbiome and Fecal
Immunoglobulin Test (FIT) to build models which could classify
individuals as having a lesion or as before surgery.\\
\textbf{Results:} Adenoma individual's bacterial microbiome were more
similar to their pre-surgery sample then those with carcinoma (P-value =
0.00198). Their change in FIT was also significantly different then
those with carcinoma (P-value = 2.15e-05). There was no significant
difference in any indivdiual OTU between samples before and after
surgery (P-value \textgreater{} 0.125). A model with a total of 37
variables was able to classify lesion (AUC = 0.847 - 0.791) while the
model to classify samples as before surgery had 33 variables (AUC = 0.79
- 0.651). The after surgery samples had a significant decrease in the
model positive probability of a lesion or before surgery sample (P-value
= 1.91e-11 and 6.72e-12). In total there were 14 OTUs that were common
to both models and the majority of these classified to commensal
bacteria (Bacteroides, Blautia, Streptococcus, and Clostridiales).\\
\textbf{Conclusions:} Our data suggests that treatment not only
significantly reduces the probability of having a colonic lesion within
our models but also causes detectable changes in the bacterial
microbiome. Further surveillance of these individuals will enable us to
determine whether models such as the ones we present could be used to
predict recurrence of colorectal cancer.

@@

\textbf{\emph{Reproducible Methods.}} A detailed and reproducible
description of how the data were processed and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005) and
the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

>>>>

\textbf{Background:} Colorectal cancer (CRC) continues to be a worldwide
health problem with early detection being used as a key component in
mitigating deaths due to the disease. Previous research suggests a link
between fecal bacterial microbiome and CRC. The overall objective of our
study was to investigate the changes in the bacterial microbiome after
surgery in patients with lesion (i.e.~adenoma or carcinoma).
Specifically, we wanted to identify what within the community was
different within those undergoing surgical removal of lesion. We also
wanted to investigate the use of the bacterial microbiome and Fecal
Immunoglobulin Test (FIT) to build models which could classify
individuals as having a lesion or as before surgery.\\
\textbf{Results:} Adenoma individual's bacterial microbiome were more
similar to their pre-surgery sample then those with carcinoma (P-value =
0.00198). Their change in FIT was also significantly different then
those with carcinoma (P-value = 2.15e-05). There was no significant
difference in any indivdiual OTU between samples before and after
surgery (P-value \textgreater{} 0.125). A model with a total of 37
variables was able to classify lesion (AUC = 0.847 - 0.791) while the
model to classify samples as before surgery had 33 variables (AUC = 0.79
- 0.651). There was a significant decrease in the model positive
probability for the after surgery sample being either a lesion or before
surgery sample (P-value = 1.91e-11 and 6.72e-12). In total there were 14
OTUs that were common to both models and the majority of these
classified to commensal bacteria (Bacteroides, Blautia, Streptococcus,
and Clostridiales).\\
\textbf{Conclusions:} Our data suggests that treatment not only
significantly reduces the probability of having a colonic lesion within
our models but also causes detectable changes in the bacterial
microbiome. Further surveillance of these individuals will enable us to
determine whether models such as the ones we present could be used to
predict recurrence of colorectal cancer.

@@

\textbf{\emph{Reproducible Methods.}} A detailed and reproducible
description of how the data were processed and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.
